\documentclass[10pt]{article}
\usepackage{amsfonts}
\usepackage{a4wide}
\usepackage{amsmath}
\thispagestyle{empty}
% for fancy 2 column lists with letters
\usepackage{multicol}
\usepackage[shortlabels]{enumitem}
\newcommand{\R}{\mathbf{R}}
\newcommand{\Z}{\mathbf{Z}}
\begin{document}


\begin{flushright} KMB,\ 16/10/17\end{flushright}

\noindent{\large \bf M1F Foundations of Analysis, Problem Sheet 2.}

\medskip
\noindent{\bf 1.} What are the following sets? Justify your answers.
%\begin{multicols}{2}
\begin{enumerate}[(a)]
\item $\bigcup_{n=0}^\infty[n,n+1)$.
\item $\bigcup_{n=1}^\infty[1/n,1]$.
\item $\bigcup_{n=1}^\infty(-n,n)$.
\item $\bigcap_{n=1}^\infty(-n,n)$.
\end{enumerate}
%\end{multicols}
\medskip
\noindent{\bf 2.} Prove that the set $(0,1)$ (that is $\{x\in\R\,:\,0<x<1\}$) has no largest element. (NB: by a ``largest element'' of a set $S$ I mean an element $x\in S$ such that $\forall y\in S, y\leq x$.)

\medskip\noindent{\bf 3.} 

(a) Prove that if $n$ is an integer and $3$ divides $n^2$ then $3$ divides~$n$.

(b) Deduce that $\sqrt{3}$ is irrational.

\medskip\noindent{\bf 4.} Are the following statements true or false? Proofs or counterexamples required.

(a) If $a$ is irrational and $b$ is irrational then $a+b$ must be irrational.

(b) If $a$ is irrational and $b$ is rational then $ab$ must be irrational.

\medskip\noindent{\bf 5.} Are the following statements true or false? Proof or counterexample required.

(a) $\forall x\in\R\,\exists y\in\R\,x+y=2.$

(b) $\exists y\in\R\,\forall x\in\R\,x+y=2.$

\medskip\noindent{\bf 6${}^*$.} Prove that $\sqrt{2}+\sqrt{6}<\sqrt{15}$ (NB you may assume the square roots exist).

\medskip\noindent{\bf 7.} Are the following numbers rational or irrational? Proofs required.

(a) $\sqrt{2}+\sqrt{3/2}$ (hint: if it were rational then its square would also be rational).

(b) $1+\sqrt{2}+\sqrt{3/2}$.

(c) $2\sqrt{18}-3\sqrt{8}$.
\end{document}

a<=b and a>=b implies a=b. Archimedean axiom.

Prove arch axiom implies for all r tyheres n with n<=r<n+1.

Prove there's no smallest positive real.

a<b implies -b<-a

1) For all x, exactly one of x > 0 or 0 > x or x = 0 is true.
(2) If x > y then x + t > y + t.
(3) If x > 0 and y > 0 then xy > 0.
(4) If x > y and y > z then x > z.
(5) For all x there is n with n > x.
